# Data Setup
Preparing data for assignment 4 by loading packages, importing data, removing NA values and and changing data class.
```{r setup, echo=FALSE}
#load packages
library(knitr)
library(psych)
library(tidyverse)
library(ggplot2)

#import dataset
nhanes <-read.csv("C:/Users/adash/Downloads/nhanes_small.csv")

#To remove NA values
nhanes <- drop_na(nhanes)

#To change character variables to factor
nhanes <- nhanes %>% 
  mutate_at(vars(3,7),list(factor))

#View the first 6 rowsdataset
kable(x = head(nhanes),caption = "Nhanes Dataset")

```

# Indexing
Learning to use **indexing lists** to print various subsets of data
```{r indexlist}
#To create a list with different number rows of the nhanes dataset

nhaneslist <- list(nhanes[1:3,], nhanes [1:6,])

#To print out these lists
nhaneslist[1]
nhaneslist[2]

```
#### Indexing lists for specific variables
Learning how to use **indexing lists** outputs to print results for *specific variables* only.
```{r}
#Using index list 1 for variable Age
nhaneslist[[1]]$Age

#Using index list 2 for variable Weight
nhaneslist[[2]]$Weight

```

# Logical Operators

#### Logical operators
Learning how to use **logical operators** to create dataframes for factor variables

*Gender* is a factor variable with two levels:

- Male
- Female
```{r logicaloperators}
#To create dataframes with factor variable: *Gender*
nhanes_male <- nhanes[nhanes$Gender == "male",]
#or
nhanes_male2 <- nhanes[nhanes$Gender != "female",]

#Both tables are the same!
#printing the first 6 rows:
kable(x = head(nhanes_male), caption = "Nhanes male") 


```

#### Combining logical operators
Learning how to combine **logical operators** using *piping*
```{r combine}
# To create a dataframe where gender = male and weight > 100:
nhanes_male3 <- nhanes%>% 
  filter(Gender == "male" & Weight > 100) 
#To view
kable(nhanes_male3)

```

# Statistical Analysis

#### Removing outliers *(+/- 2SD)*
Learning how to remove **ouliters** from the `nhanes`dataset that was loaded as `nhanes<-read.csv("C:/Users/adash/Downloads/nhanes_small.csv")`

```{r outliers}
#To remove outliers from all numeric variables:
nhanesfilter <- subset(x= nhanes, 
          subset = Weight > (mean(Weight)- 2*sd(Weight))& 
            Weight < (mean(Weight) + 2*sd(Weight))& 
            Age > (mean(Age)- 2*sd(Age)) & 
            Age < (mean(Age) + 2*sd(Age)))

# To view first 6 rows of nhanes data without outliers:
kable(x = head(nhanesfilter), caption = "Nhanes without outliers")

```
#### Summary statistics
Learning to get **summary statistics** of numeric and factor variables using *Describe/DescribeBy* functions.

```{r summarystatistics}
#To obtain summary stats for numeric values
describe(nhanesfilter[,4:6], IQR = T)

#To obtain summary stats for factor variable: Gender
describeBy(nhanesfilter, nhanesfilter$Gender)

```
We see that using`describe(nhanesfilter[,4:6], IQR = T)`gave us a the summary statistics of all our numeric variables and using `describeBy(nhanesfilter, nhanesfilter$Gender)` divided these statistics by gender.


# Scatterplots
Learning how to make **scatterplots**
```{r scatterplot}
#To make scatterplots comparing two levels of a variable:

ggplot(data = nhanesfilter, aes(x = Age, y = Weight))+ 
geom_point(aes(color = (Gender)))+
geom_point(aes(shape = (Gender)))+
geom_smooth(method ="lm", aes(color = (Gender))) 

#Scatterplot was created successfully!

```

